### 2012-12-03 paper/paper.tex
Adding the outline of the paper

% Paper for CSC580 final project.
\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{float}
%\usepackage{url}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

\begin{document}

\title{Citeopotamus}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Eriq Augustine, Aldrin Montana, Ryan Verdon}
\\
\IEEEauthorblockA{Department of Computer Science\\
Cal Poly, San Luis Obispo\\
 \textsf{\{eaugusti, amontana, rverdon\}@calpoly.edu}
}
}

\maketitle

\thispagestyle{empty}
\pagestyle{empty}

\section{Introduction}\label{sec:introduction}
%problem statement
%The whole problem of recommending citations
%Discuss the two situations we are attacking

\section{Contributions}\label{sec:contributions}
%dataset
%Citation matching methods

\section{Dataset}\label{sec:dataset}
%overview
%different components of the data
%implementation
%Metrics - #papers, #references, etc

\section{Architecture}\label{sec:architecture}

\section{Parsing}\label{sec:parsing}
%transformations
%contexts

\section{Methods}\label{sec:methods}

\section{Results}\label{sec:results}

\section{Future Work}\label{sec:future}

\section{Related Work}\label{sec:related}

\section{Conclusion}\label{sec:conclusion}

\section{Thanks}

\bibliographystyle{acm}
\bibliography{refs}

\end{document}